---
title: "Chocolate Chip Cookies"
execute:
  error: true
author: "Carson Reznicek"
output: html_document
---

## Reading In the Data

First, read in the CSV data of cookie ingredients.
Make sure that your end-result data has appropriate types for each column - these should match the types provided in the documentation in the README.md file.

```{r}
library(tidyverse)
library(skimr)
library(reticulate)

cookies <- read_csv("choc_chip_cookie_ingredients.csv")

glimpse(cookies)
```

```{python}
import pandas as pd
from skimpy import skim

cookies_py = pd.read_csv("choc_chip_cookie_ingredients.csv")
print(cookies_py.info())
```


## Exploratory Data Analysis

Exploratory data analysis is the process of getting familiar with your dataset. To get started, [this blog post](https://www.mrdbourke.com/a-gentle-introduction-to-exploratory-data-analysis/) provides a nice checklist to get you thinking:

> 1.  What question(s) are you trying to solve (or prove wrong)?
> 2.  What kind of data do you have and how do you treat different types?
> 3.  What's missing from the data and how do you deal with it?
> 4.  Where are the outliers and why should you care about them?
> 5.  How can you add, change or remove features to get more out of your data?

### Generating Questions

Generate at least 5 questions you might explore using this database of cookie ingredients.

1. How does the distribution of the most common units of measurement vary across the dataset?
2. What is the overall distribution of ingredient frequency across all unique recipes?
3. To what extent does the presence of "rare" ingredients (those in < 20 recipes) influence the variation in recipe ratings?
4. Is there a non-zero correlation between the variety of ingredients in a recipe and its average user rating?
5. What is the distribution of quantities used for "all-purpose flour" across all recipes that utilize it?


### Skimming the Data

One thing we often want to do during EDA is to examine the quality of the data - are there missing values? What quirks might exist in the dataset?

The `skimr` package in R, and the similar `skimpy` package in python (which has a much better name, in my opinion), can help provide visual summaries of the data. 

Install both packages, and read the package documentation ([R](https://cran.r-project.org/web/packages/skimr/vignettes/skimr.html), [Python](https://pypi.org/project/skimpy/)).

[Part 1] Use each package and generate summaries of your data that require the use of at least some non-default options in each package's `skim` function.


```{r}
skim(cookies) %>% 
  yank("numeric")
```

```{python}
skim(cookies_py[['Rating', 'Quantity']])
```

[Part 2] Write 1-2 sentences about what you can tell from each summary display you generate. Did you discover anything new about the data?

The R summary reveals that the Rating column has 1,010 missing values, which is a large portion of the dataset, approximately 50%, that will require careful handling during analysis. The Quantity column shows a high standard deviation relative to the mean, which likely indicates that the dataset contains a wide variety of measurement scales or units that need to be standardized.

### Generating Tables

Another useful technique for exploratory data analysis is to generate summary tables. 
You may want to use the `dplyr` package in R (`group_by` or `count` functions), as well as the `groupby` and `count` methods in Pandas. [Python example](https://sparkbyexamples.com/pandas/pandas-groupby-count-examples/), [R example](https://dplyr.tidyverse.org/reference/count.html)

[Part 1] Using R and Python, generate a table that shows what **proportion** of recipes contain each type of ingredient, for the most common 20 ingredients.

```{r}
all_recipes <- n_distinct(cookies$Recipe_Index)

top20_table <- cookies %>%
  group_by(Ingredient) %>%
  summarise(recipe_count = n_distinct(Recipe_Index)) %>%
  mutate(proportion = recipe_count / all_recipes) %>%
  arrange(desc(proportion)) %>%
  slice_head(n = 20)

print(top20_table)
```

```{python}
total_recipes = cookies_py['Recipe_Index'].nunique()

props_py = (cookies_py.groupby('Ingredient')['Recipe_Index'].nunique() / total_recipes)
print(props_py.sort_values(ascending=False).head(20))
```

[Part 2] Print out a character string that lists all of the ingredients that do not appear in at least 20 recipes.

```{r}
rare <- cookies %>%
  group_by(Ingredient) %>%
  summarise(count = n_distinct(Recipe_Index)) %>%
  filter(count < 20) %>%
  pull(Ingredient)

cat(paste(rare, collapse = ", "))
```

### Visualization

Using whatever plotting system you are comfortable with in R or python, see if you can create a couple of useful exploratory data visualizations which address one of the questions you wrote above - or another question which you've come up with as you've worked on this assignment.

[Part 1] Create at least one plot (it doesn't have to be pretty) that showcases an interesting facet of the data.

```{r}
cup_data <- cookies %>%
  filter(Unit == "cup")

top10_cups <- cup_data %>%
  count(Ingredient, sort = TRUE) %>%
  top_n(10, n) %>%
  pull(Ingredient)

cup_data %>%
  filter(Ingredient %in% top10_cups) %>%
  ggplot(aes(x = reorder(Ingredient, Quantity, FUN = median), y = Quantity)) +
  geom_boxplot(fill = "blue", outlier.color = "red", outlier.shape = 1) +
  coord_flip() + 
  labs(
    title = "Quantity Distribution for Ingredients Measured in Cups",
    subtitle = "red circles are potential outliers within the 'cup' unit",
    x = "Ingredient",
    y = "Quantity (Cups)"
  ) +
  theme_minimal()
```

[Part 2] Write 2-3 sentences about what you can learn from that plot and what directions you might want to investigate from here.

This boxplot compares the quantities of common ingredients restricted specifically to those measured in cups to ensure a valid comparison. By isolating the unit we can see that bulk ingredients like all-purpose flour have higher median quantities and wider distributions than more standard cup-based additions like walnuts. From here, I would investigate the extreme outliers (red circles) remaining in the cup data to determine if they represent 'bulk-sized' recipes or data entry errors.